\chapter{Résumé}
Ce travail de fin d’études boucle notre cursus 
universitaire pour l'obtention d'un diplôme en 
Génie Electronique à la Faculté Des Sciences de 
l'Université d'Etat d'Haïti (FDS-UEH) durant l’année 
académique 2019-2020. \par 
La gestion des données géothechniques en Haïti 
ne dispose pas d’un système informatisé centralisé. 
Dans ce mémoire, nous vous présentons une solution 
se basant sur la conception et la réalisation d'un 
tel système. \par 
La totalité des codes est disponible en ligne sur 
GitHub :https://github.com/geotech. L’application est 
hébergée par AWS à l’adresse: https://geotech.ht \par 
Mots-clés : Données spatiales, Croquis.

\paragraph{Abstract}
\par 
The sketch is an old form of communication that was 
used to visualize, share and store information. 
Despite its proven expressiveness, it has not yet 
become a used modality in the interaction between 
humans and computer systems. Geographic Information 
Systems (GIS) have special needs for such advanced 
forms of interaction, because they involve complex and
heterogeneous data structures, which are often 
difficult to describe by the text or by predicates 
based on attributes. The objective of this work is to 
show the usefulness of the sketch in retrieval spatial 
information in a particular case which is the 
geographic data. In this context, a prototype 
application of sketch query system is implemented 
in order to verify the concepts and theories 
developed. \par 
Keywords : Spatial data, Sketch.